---
title: "Tuition Trends in Maryland and Virginia Colleges and Universities"
output: html_notebook
Name: Karyme Bravo
---
# Notes
# 1. Data Source and Description
# The assignment relates to a public dataset with the tuition info.
# Tuition names, tuition levels, and the year information
# Verifying where the CSV comes from and documentation to be place in my folder

# I will add the links in Rmd as the dataset is broken down by school and year

# 2. The data source itself
# It would include the tuition info for Maryland and Virginia across their colleges and universities
# The links that were provided from ELMS that are an external site.


# 3. Notes on the data source and time period
# Each row contains one school in that one year.
# It would perhaps cover more than a decade (like 10 to 20 years)
# Covers both public and private colleges

# 4. Columns that will be important for the analysis
# When I would load the data, it would have columns like:
# institution_name
# year
# State (MD or VA)
# tuition_out_of_state
# tuition_in_state
# type (whether public or private)

# 5. What would be missing
# The enrollment data
# Details on public and private institutions unless it is included seperately

# 6. Descriptive Stats (min, med, max)
```{r}
tuition <- read.csv("https://raw.githubusercontent.com/wellsdata/Jour405_student_repo_fall_2025/refs/heads/main/project/college_scorecard_2021_2024.csv")
```
# Maryland
```{r}
md <- tuition[tuition$state == "MD",]
mean(md$X2023_24_tuitionfee_in, na.rm = TRUE)
median(md$X2023_24_tuitionfee_in, na.rm = TRUE)
min(md$X2023_24_tuitionfee_in, na.rm = TRUE)
max(md$X2023_24_tuitionfee_in, na.arm = TRUE)
sd(md$X2023_24_tuitionfee_in, na.rm = TRUE)
```

# 7. Propose hypothesis newsworthy
# Hypothesis 1: The tuition has increased in at a faster rate in Virginia than in Maryland
# Hypothesis 2: Out of state tuition has increased and grown rapidly than in state tuition
# Hypothesis 3: There are certain years that suddenly show an increase that can relate to the economy 
# Hypothesis 4: Private colleges are the reason for the tuition increase.

# 8. Questions
# Should I compare median across the institutions and avoid private schools?
# Should I adjust tuition increase that compares to the year?

Part 2: Loading the data
Load Data and inspection
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
tuition <- read.csv("https://raw.githubusercontent.com/wellsdata/Jour405_student_repo_fall_2025/refs/heads/main/project/college_scorecard_2021_2024.csv") |>
  arrange(state) # Sort by state
```

Subset and Descriptive Stats of MD and VA
#Maryland Stats
```{r}
md <- tuition[tuition$state == "MD",]
mean(md$X2023_24_tuitionfee_in, na.rm = TRUE)
median(md$X2023_24_tuitionfee_in, na.rm = TRUE)
min(md$X2023_24_tuitionfee_in, na.rm = TRUE)
max(md$X2023_24_tuitionfee_in, na.arm = TRUE)
sd(md$X2023_24_tuitionfee_in, na.rm = TRUE)
```

#Virginia Stats
```{r}
va <- tuition[tuition$state == "VA",]
mean(va$X2023_24_tuitionfee_in, na.rm = TRUE)
median(va$X2023_24_tuitionfee_in, na.rm = TRUE)
min(va$X2023_24_tuitionfee_in, na.rm = TRUE)
max(va$X2023_24_tuitionfee_in, na.arm = TRUE)
sd(va$X2023_24_tuitionfee_in, na.rm = TRUE)
```
Z-score
```{r}
```

T-test
```{r}
# Period of the timeline
t_test_states <- t.test(tuition_in ~ state, data = tuition)

# Display results
t_test_states

# Visualization of Data
state_summary <- tuition |>
  group_by(state) |>
  summarize(
    mean_tuition = mean(tuition_in, na.rm = TRUE),
    sd_tuition = sd(tuition_in, na.rm = TRUE),
    schools = n(),
    se_tuition = sd_tuition / sqrt(schools),
    margin_error = qt(0.975, df = schools - 1) * se_tuition, 
    # look at how we calculated this above
    ci_lower = mean_tuition - margin_error,
    ci_upper = mean_tuition + margin_error
  )
```

Data visualization
```{r}
```




