---
title: "Tuition Trends in Maryland and Virginia Colleges and Universities"
output: html_notebook
Name: Karyme Bravo
---
# Notes
# 1. Data Source and Description
# The assignment relates to a public dataset with the tuition info.
# Tuition names, tuition levels, and the year information
# Verifying where the CSV comes from and documentation to be place in my folder

# I will add the links in Rmd as the dataset is broken down by school and year

# 2. The data source itself
# It would include the tuition info for Maryland and Virginia across their colleges and universities
# The links that were provided from ELMS that are an external site.


# 3. Notes on the data source and time period
# Each row contains one school in that one year.
# It would perhaps cover more than a decade (like 10 to 20 years)
# Covers both public and private colleges

# 4. Columns that will be important for the analysis
# When I would load the data, it would have columns like:
# institution_name
# year
# State (MD or VA)
# tuition_out_of_state
# tuition_in_state
# type (whether public or private)

# 5. What would be missing
# The enrollment data
# Details on public and private institutions unless it is included seperately

# 6. Descriptive Stats (min, med, max)
```{r}
tuition <- read.csv("https://raw.githubusercontent.com/wellsdata/Jour405_student_repo_fall_2025/refs/heads/main/project/college_scorecard_2021_2024.csv")
```
# Maryland
```{r}
md <- tuition[tuition$state == "MD",]
mean(md$X2023_24_tuitionfee_in, na.rm = TRUE)
median(md$X2023_24_tuitionfee_in, na.rm = TRUE)
min(md$X2023_24_tuitionfee_in, na.rm = TRUE)
max(md$X2023_24_tuitionfee_in, na.arm = TRUE)

```
# 7. Propose hypothesis newsworthy
# Hypothesis 1: The tuition has increased in at a faster rate in Virginia than in Maryland
# Hypothesis 2: Out of state tuition has increased and grown rapidly than in state tuition
# Hypothesis 3: There are certain years that suddenly show an increase that can relate to the economy 
# Hypothesis 4: Private colleges are the reason for the tuition increase.

# 8. Questions
# Should I compare median across the institutions and avoid private schools?
# Should I adjust tuition increase that compares to the year?
