---
title: "Tuition Trends in Maryland and Virginia Colleges and Universities"
output: html_notebook
Name: Karyme Bravo
---
# Notes
# 1. Data Source and Description
# The assignment relates to a public dataset with the tuition info.
# Tuition names, tuition levels, and the year information
# Verifying where the CSV comes from and documentation to be place in my folder

# I will add the links in Rmd as the dataset is broken down by school and year

# 2. The data source itself
# It would include the tuition info for Maryland and Virginia across their colleges and universities
# The links that were provided from ELMS that are an external site.


# 3. Notes on the data source and time period
# Each row contains one school in that one year.
# It would perhaps cover more than a decade (like 10 to 20 years)
# Covers both public and private colleges

# 4. Columns that will be important for the analysis
# When I would load the data, it would have columns like:
# institution_name
# year
# State (MD or VA)
# tuition_out_of_state
# tuition_in_state
# type (whether public or private)

# 5. What would be missing
# The enrollment data
# Details on public and private institutions unless it is included seperately

# 6. Descriptive Stats (min, med, max)
```{r}
tuition <- read.csv("https://raw.githubusercontent.com/wellsdata/Jour405_student_repo_fall_2025/refs/heads/main/project/college_scorecard_2021_2024.csv")
```

# Maryland
```{r}
md <- tuition[tuition$state == "MD",]
mean(md$X2023_24_tuitionfee_in, na.rm = TRUE)
median(md$X2023_24_tuitionfee_in, na.rm = TRUE)
min(md$X2023_24_tuitionfee_in, na.rm = TRUE)
max(md$X2023_24_tuitionfee_in, na.rm = TRUE)
sd(md$X2023_24_tuitionfee_in, na.rm = TRUE)
```

# 7. Propose hypothesis newsworthy
# Hypothesis 1: The tuition has increased in at a faster rate in Virginia than in Maryland
# Hypothesis 2: Out of state tuition has increased and grown rapidly than in state tuition
# Hypothesis 3: There are certain years that suddenly show an increase that can relate to the economy 
# Hypothesis 4: Private colleges are the reason for the tuition increase.

# 8. Questions
# Should I compare median across the institutions and avoid private schools?
# Should I adjust tuition increase that compares to the year?

Part 2: Loading the data
# Load Data and inspection
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
tuition <- read.csv("https://raw.githubusercontent.com/wellsdata/Jour405_student_repo_fall_2025/refs/heads/main/project/college_scorecard_2021_2024.csv") |>
  arrange(state) # Sort by state
```

Subset and Descriptive Stats of MD and VA
#Maryland Stats
```{r}
md <- tuition[tuition$state == "MD",]
mean(md$X2023_24_tuitionfee_in, na.rm = TRUE)
median(md$X2023_24_tuitionfee_in, na.rm = TRUE)
min(md$X2023_24_tuitionfee_in, na.rm = TRUE)
max(md$X2023_24_tuitionfee_in, na.rm = TRUE)
sd(md$X2023_24_tuitionfee_in, na.rm = TRUE)
```
# Data Explanation for Maryland
This is Maryland institutions that resulted in 93 schools. The average in-state tuition for Maryland colleges is at 18,035.2 which represents the average tuition cost across all Maryland institutions in the dataset. The median is 11,100 which means half of Maryland schools charge less than this amount and half charge more. This indicates that higher priced schools may be pulling the mean upward.
Additionally, the minimum tuition is 3,314 while the maximum is 63,340, which shows a very wide range in tuition cots across the state of Maryland instutions. Lastly, the standard deviation is 17,663.76 which shows a substantial variation in tuition prices and highlighting the large difference between lower-cost and higher-cost schools.


# Virginia Stats
```{r}
va <- tuition[tuition$state == "VA",]
mean(va$X2023_24_tuitionfee_in, na.rm = TRUE)
median(va$X2023_24_tuitionfee_in, na.rm = TRUE)
min(va$X2023_24_tuitionfee_in, na.rm = TRUE)
max(va$X2023_24_tuitionfee_in, na.rm = TRUE)
sd(va$X2023_24_tuitionfee_in, na.rm = TRUE)
```
# Data Explanation for Virginia
In the Virginia dataset, there are about 172 Virginia colleges and universities. The mean in-state tuition is 17,866.68, which is slightly lower than Maryland's average tuition. As the median is 15,200 which is closer to the mean than in Maryland that indicates a more evenly distributed tuition structure. Whereas, the minimum is about 4,835 an the maximum is 64,525, as this shows Virginia has a vast ride spread of tuition prices. Additionally, the standard deviation is 12,457.41 which is lower than Maryland's as this shows the tuition costs in Virginia is more consistent in the institutions.


# Means
```{r}
md_mean <- mean(md$X2023_24_tuitionfee_in, na.rm = TRUE)
va_mean <- mean(va$X2023_24_tuitionfee_in, na.rm = TRUE)
```
This explains the average tuition for each state in varaibles that would be used as a reference and it helps compare the overall tuition costs between Maryland and Virginia.


# Z-score
```{r}
overall_mean <- mean(tuition$X2023_24_tuitionfee_in, na.rm = TRUE)
overall_sd <- sd(tuition$X2023_24_tuitionfee_in, na.rm = TRUE)

```
This shows the overall standard deviation for the entire dataset itself. These numbers will be used to calculate the z-score to compare each schools of their average costs for both states.



# T-test
```{r}
# Period of the timeline
t_test_states <- t.test(X2023_24_tuitionfee_in ~ state, data = tuition)

# Display results
t_test_states
```
I ran a t-test to indicate whether if there is a difference in-state tuition between Maryland and Virginia as it helps one of my hypothesis: as whether tuition is higher in one state than the other, which Maryland seems to have the higher cost.



# Data Visualization
```{r}
ggplot(tuition, aes(x = state, y = X2023_24_tuitionfee_in)) +
  geom_boxplot() +
  labs(
    title = "Distribution of In-State Tuition by State",
    x = "State",
    y = "In-State Tuition"
  )
theme_minimal()
```
This boxplot shows the distribution of the tuition prices for Maryland and Virginia. We see that Maryland's boxplot is more likely to show a more higher costs due to private schools, while Virginia is more even.



# Final Draft Story
Title: "Maryland vs. Virginia: Who Pays More for College?"

Students in Maryland and Virginia states are facing a difference in tuition. For the 2023-24 school year, as Maryland's average in-state tuition is $18,035, which is slightly higher than Virginia's in-state tuition which is at $17,867. However, the median is $11,100 in Maryland and $15,200 in Virginia as it displays how expensive private schools in Maryland push the average up, whole Virginia schools stays consistent and closer to the middle. Tuition in Maryland ranges from $3,314 to $63,340 with a standard deviation of $17,664. As Virginia ranges from $4,835 to $64,525, with a standard deviation of $12,457.

There are about 90 schools that have missing tuition data and enrollment numbers that are not included, as the averages do not reflect the student. Additionally, private schools are skewing the Maryland numbers.

# Reporting Plan
The story will use boxplot and descriptive statistics that would visualize these patterns. It would examine whether public or private schools drive those tuition trends. The reporting would show how the differences between the two states can affect students affordability and their choices which can help inform families what to expect on the cost when applying for colleges; whether it is in the region or if it is in state.







