---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
#install.packages('tidyverse')
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities |>
  mutate(
    violent_rate_2019 = (violent2019 / pop2019) * 100000,
    property_rate_2019 = (property2019 / pop2019) *100000,
    violent_rate_2020 = (violent2020 / pop2020) * 100000,
    property_rate_2020 = (property2020 / pop2020) * 100000
  )
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

There are some cities that have the same highest violent and property crime rates in 2019 and 2020. This can indicate that there are different types of crime trends. The story is there are a few cities that drive most of the rates year to year that can be affected in different states.

### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |>
  mutate(
    violent_rate_pct_change = ((violent_rate_2020 - violent_rate_2019) / violent_rate_2019) *100,
    property_rate_pct_change = ((property_rate_2020 - property_rate_2019) / property_rate_2019) * 100
  )
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

Some cities show a shift year to year but some have a large increase in violent or property crime rates. This shows that in Maryland cities, are in suburban areas.

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Violent crime rates increased in Maryland cities from 2019 and 2020 as property crime decreased.

The data showed that crime trends are inconsistent across Maryland as most cities saw a change, but a few showed increase in violent crime. Many communities see a shift as it indicates for residents of where they may feel safe which allows local leaders to focus on what they can do to make it safer.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.

### Task 10: Calculate percentile using ntile

```{r}
md_cities_percentiles <- md_cities_change |>
  mutate(percentile = ntile(violent_rate_2020, 100))
```

```{r}
md_cities_change |>
summarize(md_cities_change, iqr_violent_2020 = IQR(violent_rate_2020, na.rm = TRUE))
```

This does not change my story idea because the IQR confirmed that most of the cities within the middle range of violent crime rate. Whereas smaller group of cities can be both high or lower.
