---
title: "Oct 1 Activities: M&Ms; Survey; Sampling"
date: 10-1-2025
---

All three class activities are on this .rmd notebook

Activity #1: Activity: M&Ms - Central Limit Theorem

Working in groups, each group will get a handful of candy from the bag, and then count the number of red and black candies. Log those in this Google Sheet 

https://docs.google.com/spreadsheets/d/1KOeoVTq2W7L2T9sWjwh5U_2f59GsTeAylVGD53rNWww/edit?usp=sharing

(give your group a nice name).

Take a look at the values that we all get - what do you notice about the patterns? This is an exercise about the central limit theorem. 

The central limit theorem states that the sampling distribution of the mean of any independent, random variable will be normal or nearly normal, if the sample size is large enough.

Then discuss and try to answer these questions:

--How close is your group's mean to the true proportion of 0.5?
--Why do different groups get different results?
--What happens to individual extreme results when we combine all groups?
--How does this relate to polls and surveys in real life?

Write your answer below. And then you may eat the M&Ms.

**A:**
- It is far away from being 0.5.
- Because each bag og M&M's have different proportion of colors; not all colors are the same so everyone gets a different result.
- All results are are around 80% to 92% which is far.
- In real life, we would have to interpret how the data we found would relate to our story. Sometimes the result may not be what we had hoped which data can be unpredictable and we may have to adjust our story to the data we had found, or we find more data that relates to our story and can use some of what we had found earlier.

------------------------------------------------------------------------------

# Activity #2:

Belonging Survey

Working in pairs, take a look back at the main findings from UMD's Belonging Survey:

https://president.umd.edu/internal/belonging


Specifically, focus on the "Sample Description" section and the Appendix. Discuss the differences and between the population and the sample, and what the document says about how the survey attempted to deal with them. Are there any decisions described in the document that raise questions in your minds? Detail the differences you see and your understanding of the methods they employed (it's ok if you don't understand all of them), and add your questions. You should discuss this as a pair but submit individually.

**A:**
It seems the results targeted the entire campus community but only portions of students, faculty members and staff. It seemed the population was around 53,450 as it still raises the question if there was an over-representation of certain groups? How is it that they were able to get bias responses since it was a choice?
------------------------------------------------------------------------------

# TASK 3: SAMPLING

Based on the class sampling demo (available on Elms), sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

**A**
```{r}
library(tidyverse)

```


### Task 2: Load 1,000 rows of data with the ages of people from 18-80

**A**
```{r}
set.seed(123)
population <- tibble(
  age = sample(18:80, size = 1000, replace = TRUE)
)

head(population)

```

### Task 3: Sample 100 rows of data from the dataset
**A**
```{r}
sample_data <- sample_n(population, 100)
head(sample_data)

```


### Task 4: Calculate the mean age and standard deviation of the sample
**A**
```{r}
sample_data |> summarize(mean = mean(age), median = median(age), sd = sd(age))

```


### Task 5: Calculate the mean age and standard deviation of the population
**A**
```{r}
population |> summarize(mean = mean(age), median = median(age), sd = sd(age))

```


### Task 6: Compare the sample and population statistics
**A**
```{r}
sample_stats <- sample_data |>
  summarize(mean = mean(age), sd = sd (age))
population_stats <- population |> summarize(mean = mean(age), sd = sd(age))

sample_stats
population_stats

```

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

**A**:
They are some what similar but not all the way, but with smaller samples, it gives it becomes less reliable as there is not much information that can be useful.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
